---
title: "Lab 2"
author: "Corinna Hong"
date: "January 16, 2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r, load packages, echo = FALSE, message = FALSE, warning = FALSE}

library(tidyverse)
library(janitor)
library(naniar)
library(VIM)
library(skimr)
library(ggfortify)
library(here)

```


##1. Get some data
```{r, read in data, message = FALSE}

ca_pb <- read_csv(here("data","ca_pollution_burden.csv"))
ca_dem <- read_csv(here("data","ca_census_demographics_2010.csv"))

```


##2. Clean and wrangle data
```{r, message = FALSE}

ca_pb_nopct <- ca_pb %>% 
  clean_names() %>%
  select(-contains("perc")) %>% # - means remove contains() means anything that contains what is in quotes
  select(-contains("pctl")) %>% 
  select(-latitude, - longitude) # no quotes are for column names


ca_dem_clean <- ca_dem %>% 
  clean_names()

```


##3. Pollution burden indicators only
```{r, message = FALSE}

ca_pb_subest <- ca_pb_nopct %>% 
  select(ozone:solid_waste,asthma:housing_burden)

```

Doing PCA
```{r, eval = FALSE}

# eval = FALSE up top lets us keep failed code for our record w/o running into problems with knitting

pb_pca <- prcomp(ca_pb_subest, scale = TRUE) # This gives an error because there are missings!

```

*Aside: Explore missingness*
```{r}

# summary(ca_pb_subest) can tell you how many NA's there are in the data . In our case, relatively low proportion. Removing them would be ok
gg_miss_var(ca_pb_subest) # graph to show you how many NA's for each variable 
matrixplot(ca_pb_subest) # Red areas are missing data and can also help visualize correlation

```

Remove (complete case deletion) any observation with ANY NA values:
```{r, results = FALSE}

ca_pb_nona <- ca_pb_subest %>% 
  drop_na()

skim(ca_pb_nona) # summarizes in a nicer way than summary()

```

Doing PCA II
```{r, warning = FALSE}

my_ca_pca <- prcomp(ca_pb_nona, scale = TRUE)

my_ca_pca # Gives loading for every PC
summary(my_ca_pca) # Proportion of variance explained by each PC

# biplot(my_ca_pca) that was very messy

# Actually, let's use autoplot

my_biplot <- autoplot(my_ca_pca,
                      colour = NA,
                      loadings.label = TRUE,
                      loadings.label.size = 3,
                      loadings.label.colour = "black",
                      loadings.label.repel = TRUE) + # forces labels to "repel" each other a little
  theme_minimal()

my_biplot

```


##4. PCA for pollution burden + demographics

join together the ca_dem_clean with ca_pb_nopct
```{r}

ca_df <- ca_dem_clean %>%
  inner_join(ca_pb_nopct, by = c("census_tract_number" = "census_tract"))

# Now drop NAs by complete case deletion

ca_nona <- ca_df %>%
  drop_na()

```

make a subset of ca_nona that only includes:\
-white percent\
-elderly_65_percent\
-pm2_5\
-pesticides\
-traffic\
-asthma\
-cardiovascular_disease\
-poverty\

```{r}

ca_subset <- ca_nona %>%
  select(white_percent, elderly_65_percent, pm2_5, pesticides, traffic, asthma, cardiovascular_disease, poverty)

```

PCA
```{r, warning = FALSE}

my_dem_pca <- prcomp(ca_subset, scale = TRUE)

my_dem_biplot <- autoplot(my_dem_pca,
                      colour = NA,
                      loadings.label = TRUE,
                      loadings.label.size = 3,
                      loadings.label.colour = "black",
                      loadings.label.repel = TRUE) + # forces labels to "repel" each other a little
  scale_y_continuous(limits = c(-0.05,0.05)) +
  theme_minimal()

my_dem_biplot

```

